---
title: "Summary statistics, tables and plots"
author: "Adrian Barnett"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "reports") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, error=FALSE, comment='', dpi=400)
library(dplyr)
library(stringr)
library(tableone) # for giant baseline table
library(binom) # for CIs
library(xtable) # for latex tables
library(flextable)
library(ggplot2)
g.theme = theme_bw() + theme(panel.grid.minor = element_blank())

# for x and y matrices/vector
load('data/3_plus_experience.RData') # from 3_add_author_experience.R
# remove missing country, subjects, and author experience to match analysis data; added late exclusion for Formal Comments
data_for_table = dplyr::filter(data, !is.na(country), !is.na(author_papers), lengths(subjects)>1, type!='Formal Comment')
n_rows = nrow(data_for_table)
data_for_table = mutate(data_for_table, year = as.numeric(format(published, '%Y'))) # add year published

# to get X matrix and Y
source('4_data_prepare.R')
```

## Top ten countries

```{r top_ten_countries}
## find top ten countries across all results
# Remove double countries per article:
countries = lapply(data_for_table$country, unique)
countries = unlist(countries)
n_country = length(unique(countries))
tab = table(countries)
top_ten = names(tab[order(-tab)][1:10]) # get the top ten countries

# now apply to x matrix and y vector to get top ten per group
cpattern = str_replace_all(paste(paste('country_', top_ten, sep=''), collapse='|'),' ','_') # make country pattern to search for
x_columns_to_use = str_detect(colnames(x), pattern = cpattern)
if(sum(x_columns_to_use) != 10){cat('Missing countries')}
x_country = x[,x_columns_to_use]

## results for each group
# review open
tab1 = colSums(x_country[y==TRUE,]) %>%
  as.data.frame() 
names(tab1) = 'count'
tab1 = mutate(tab1, 
              country = rownames(tab1),
              country = str_remove(country, '^country_'),
              country = str_replace_all(country, '_', ' '),
              percent = round(100*count/sum(y==TRUE)),
              yes = paste(count, ' (', percent, ')', sep='')) %>%
  select(-count, -percent)
rownames(tab1) = NULL
# review not open
tab2 = colSums(x_country[y==FALSE,]) %>%
  as.data.frame() 
names(tab2) = 'count'
tab2 = mutate(tab2, 
              country = rownames(tab2),
              country = str_remove(country, '^country_'),
              country = str_replace_all(country, '_', ' '),
              percent = round(100*count/sum(y==FALSE)),
              no = paste(count, ' (', percent, ')', sep=''))%>%
  select(-percent)
rownames(tab2) = NULL
#
tab = full_join(tab2, tab1, by='country') %>%
  arrange(desc(count)) %>%
  rename('Country' = 'country') %>%
  select(-count)
ftab = flextable(tab) %>%
  theme_box() %>%
  autofit()
ftab
# export to latex
print(xtable(tab), include.rownames=FALSE, hline.after=FALSE, file = "results/5_countries.tex")
```

There were `r n_country` countries in total. 
The table shows the authors' countries where each country is counted once per article. The percentages are over all `r format(n_rows,big.mark=',')` articles.

## Top ten subjects

```{r top_ten_subjects}
## find top ten subjects across all results
# Remove double subjects per article:
subjects = lapply(data_for_table$subjects, unique)
subjects = unlist(subjects)
n_subject = length(unique(subjects))
tab = table(subjects)
top_ten = names(tab[order(-tab)][1:10]) # get the top ten subjects

# now apply to x matrix and y vector to get top ten per group
cpattern = str_replace_all(paste(paste('subject_', top_ten, sep=''), collapse='|'),' ','_') # make subject pattern to search for
x_columns_to_use = str_detect(colnames(x), pattern = cpattern)
if(sum(x_columns_to_use) != 10){cat('Missing subjects')}
x_subject = x[,x_columns_to_use]

## results for each group
# review open
tab1 = colSums(x_subject[y==TRUE,]) %>%
  as.data.frame() 
names(tab1) = 'count'
tab1 = mutate(tab1, 
              subject = rownames(tab1),
              subject = str_remove(subject, '^subject_'),
              subject = str_replace_all(subject, '_', ' '),
              percent = round(100*count/sum(y==TRUE)),
              yes = paste(count, ' (', percent, ')', sep='')) %>%
  select(-count, -percent)
rownames(tab1) = NULL
# review not open
tab2 = colSums(x_subject[y==FALSE,]) %>%
  as.data.frame() 
names(tab2) = 'count'
tab2 = mutate(tab2, 
              subject = rownames(tab2),
              subject = str_remove(subject, '^subject_'),
              subject = str_replace_all(subject, '_', ' '),
              percent = round(100*count/sum(y==FALSE)),
              no = paste(count, ' (', percent, ')', sep=''))%>%
  select(-percent)
rownames(tab2) = NULL
#
tab = full_join(tab2, tab1, by='subject') %>%
  arrange(desc(count)) %>%
  rename('subject' = 'subject') %>%
  select(-count)
ftab = flextable(tab) %>%
  theme_box() %>%
  autofit()
ftab
# export to latex
print(xtable(tab), include.rownames=FALSE, hline.after=FALSE, file = "results/5_subjects.tex")
```

There were `r n_subject` subjects in total. 
The percentages are over all `r format(n_rows,big.mark=',')` articles.

## Top ten email domains

```{r top_domains}
## find top ten domains across all results
# Remove double domains per article:
domains = lapply(data_for_table$edomain, unique)
domains = unlist(domains)
n_domain = length(unique(domains))
tab = table(domains)
top_ten = names(tab[order(-tab)][1:10]) # get the top ten domains

# now apply to x matrix and y vector to get top ten per group
cpattern = str_replace_all(paste(paste('domain_', top_ten, sep=''), collapse='|'),' ','_') # make domain pattern to search for
x_columns_to_use = str_detect(colnames(x), pattern = cpattern)
if(sum(x_columns_to_use) != 10){cat('Missing domains')}
x_domain = x[,x_columns_to_use]

## results for each group
# review open
tab1 = colSums(x_domain[y==TRUE,]) %>%
  as.data.frame() 
names(tab1) = 'count'
tab1 = mutate(tab1, 
              domain = rownames(tab1),
              domain = str_remove(domain, '^domain_'),
              domain = str_replace_all(domain, '_', ' '),
              percent = round(100*count/sum(y==TRUE)),
              yes = paste(count, ' (', percent, ')', sep='')) %>%
  select(-count, -percent)
rownames(tab1) = NULL
# review not open
tab2 = colSums(x_domain[y==FALSE,]) %>%
  as.data.frame() 
names(tab2) = 'count'
tab2 = mutate(tab2, 
              domain = rownames(tab2),
              domain = str_remove(domain, '^domain_'),
              domain = str_replace_all(domain, '_', ' '),
              percent = round(100*count/sum(y==FALSE)),
              no = paste(count, ' (', percent, ')', sep=''))%>%
  select(-percent)
rownames(tab2) = NULL
#
tab = full_join(tab2, tab1, by='domain') %>%
  arrange(desc(count)) %>%
  rename('domain' = 'domain') %>%
  select(-count)
ftab = flextable(tab) %>%
  theme_box() %>%
  autofit()
ftab
# export to latex
print(xtable(tab), include.rownames=FALSE, hline.after=FALSE, file = "results/5_domain.tex")
```

Some papers had multiple emails. 
There were `r n_domain` domains in total. 
The percentages are over all `r format(n_rows,big.mark=',')` articles.

## Top ten funders

```{r top_funders}
## find top ten funders across all results
# Remove double funders per article:
funders = lapply(data_for_table$funder_number_consolidated, unique)
funders = unlist(funders)
funders[funders=='100010269'] = '100004440' # combine two Wellcome
n_funder = length(unique(funders))
tab = table(funders)
top_ten = names(tab[order(-tab)][1:10]) # get the top ten funders

# now apply to x matrix and y vector to get top ten per group
cpattern = str_replace_all(paste(paste('funder_', top_ten, sep=''), collapse='$|'),' ','_') # make funder pattern to search for
x_columns_to_use = str_detect(colnames(x), pattern = cpattern)
if(sum(x_columns_to_use) != 10){cat('Missing funders\n')}
x_funder = x[,x_columns_to_use]

## results for each group
# review open
tab1 = colSums(x_funder[y==TRUE,]) %>%
  as.data.frame() 
names(tab1) = 'count'
tab1 = mutate(tab1, 
              funder = rownames(tab1),
              funder = str_remove(funder, '^funder_'),
              funder = str_replace_all(funder, '_', ' '),
              percent = round(100*count/sum(y==TRUE)),
              yes = paste(count, ' (', percent, ')', sep='')) %>%
  select(-count, -percent)
rownames(tab1) = NULL
# review not open
tab2 = colSums(x_funder[y==FALSE,]) %>%
  as.data.frame() 
names(tab2) = 'count'
tab2 = mutate(tab2, 
              funder = rownames(tab2),
              funder = str_remove(funder, '^funder_'),
              funder = str_replace_all(funder, '_', ' '),
              percent = round(100*count/sum(y==FALSE)),
              no = paste(count, ' (', percent, ')', sep=''))%>%
  select(-percent)
rownames(tab2) = NULL
#
tab = full_join(tab2, tab1, by='funder') %>%
  arrange(desc(count)) %>%
  rename('funder' = 'funder') %>%
  select(-count) %>%
  mutate(funder = case_when( # replace numbers with names
    funder == '501100000781' ~ 'European Research Council',
    funder == '501100000265' ~ 'Medical Research Council',
    funder == '501100001809' ~ 'National Natural Science Foundation of China',
    funder == '100000001' ~ 'National Science Foundation',
    funder == '100000002' ~ 'National Institutes of Health',
    funder == '100004440' ~ 'Wellcome', 
    funder == '501100001321' ~ 'National Research Foundation, South Africa',
    funder == '100000060' ~ 'National Institute of Allergy and Infectious Diseases',
    funder == '100010002' ~ 'Ministry of Education, China',
    funder == '501100002701' ~ 'Ministry of Education, Republic of Korea',
    funder == '501100001691' ~ "Japan Society for the Promotion of Science",
    TRUE ~ as.character(funder)
  ))
ftab = flextable(tab) %>%
  theme_box() %>%
  autofit()
ftab
# export to latex
print(xtable(tab), include.rownames=FALSE, hline.after=FALSE, file = "results/5_funder.tex")
```

Some papers had multiple funders. 
There were `r n_funder` funders in total, hence we only show the top ten.

## Summary table

```{r summary}
# Create a TableOne object
vars = c('review_available','year','n_authors','p_orcid','author_papers','time_between','journal','type','citations','retracted') # 
n_vars = c('year','n_authors','p_orcid','time_between','citations','author_papers') # number variables
tab2 <- CreateTableOne(vars = vars, 
                       data = data_for_table, 
                       strata = 'review_available',
                       factorVars = NULL, 
                       test = FALSE, 
                       includeNA = TRUE,
                       addOverall = FALSE) # no total column
print(tab2, # print version for Word document
      thenonnormal = n_vars, # use median instead of mean
      catDigits = 1, # no longer allows 0 
      contDigits = 1, 
      explain = TRUE, 
      showAllLevels = TRUE,
      formatOptions = list(big.mark = ""))
# export to latex
m2 = data.frame(print(tab2, 
                      printToggle = FALSE,
                      nonnormal = n_vars, # use median instead of mean
                      catDigits = 1, # no longer allows 0 
                      contDigits = 1, 
                      explain = TRUE, 
                      showAllLevels = TRUE,
                      formatOptions = list(big.mark = ""))) %>%
  tibble::rownames_to_column(var = "var")
print(xtable(m2), include.rownames=FALSE, hline.after=FALSE, file = "results/5_big_table.tex")
```

## Openness over time

```{r time}
summary = group_by(data_for_table, year) %>%
  summarise(open = sum(review_available), 
            n = n(),
            p = open/n,
            lower = binom.confint(open, n, methods = 'asymptotic')$lower,
            upper = binom.confint(open, n, methods = 'asymptotic')$upper)
gplot = ggplot(data = summary, aes(x=year, y=p, ymin=lower,ymax=upper))+
  geom_point()+
  geom_line()+
  geom_ribbon(alpha=0.5)+
  xlab('Year')+
  ylab("Proportion choosing open peer review")+
  g.theme
gplot
```

The plot shows the yearly proportion and 95% confidence interval.

There was an increase over time in the proportion of authors choosing open review.